---
title: 'Homework #2: SDGB-7844'
author: "Matthew Fein"
fordham_id_number: "A16797353"
fordham_email: "mfein@fordham.edu"
date: "10/31/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Choose 5 problems below to complete.  Each question below comes from our text, **Introduction to Scientific Programming and Simulation with R**.  Please comment your code where appropriate.

### Chapter 3 - Exercise 1

```{r}
# input
x.values <- seq(-2, 2, by = 0.1)
# for each x calculate y
n <- length(x.values)

y.values <- rep(0, n)
for (i in 1:n) {
x <- x.values[i]

if (x < 0) {
y<- -x**3
} else if (x > 0 & x <1) {
y <- x**2
} else {
  y <- sqrt(x)
}

y.values[i] <- y
}
# output
plot(x.values, y.values, type = "l")

```

### Chapter 3 - Exercise 2

```{r}
#x <- readline(prompt="Enter a value for x: ")
#n <- readline(prompt="Enter a value for n: ")
x<- 3
n<- 5

h<-as.integer(c(1, rep(x, times = n)))

for(item in 2:length(h)){
  
  # print value of item
  h[item] <-  h[item]**(item-1)
  print( h[item])
  
 # l
 print( h )

}
```
print(sum(h))

### Chapter 3 - Exercise 4


```{r}
#x <- readline(prompt="Enter a value for x: ")
#n <- readline(prompt="Enter a value for n: ")
x<- 3
n <- 5
h<-as.integer(c(1, rep(x, times = n)))

i <- 2
while(i < length(h)){
  
  # print value of item
  h[i] <-  h[i]**(i-1)
  print( h[i])
  
 # l
 print( h )
 i<- i+1

}
```
print(sum(h))

```

### Chapter 3 - Exercise 6


```{r}
 

```

### Chapter 3 - Exercise 7


```{r}
x <- seq(0:12)
 
sum3rd <- sum((x[i:length(x)%%3==0]))
```


### Chapter 3 - Exercise 12

```{r}
x <- sum(ceiling(6*runif(2)))
print(sprintf("Roll is %s.", x))

if (x==7|x==11) {
  print("You win!")

} else if (x == 2|x == 3|x == 12) {
  print("You lose!")
  
} else {

  print(sprintf("The point is on '%s', shoot again.",x))  
  
  x2<-0
  while(x2 != x){
  x2 <- sum(ceiling(6*runif(2)))
  if (x2==7) {
    print("Craps you lose!")
    break
  } else{
    print(sprintf("%s, Shoot again.", x2))
    x2 <- sum(ceiling(6*runif(2)))
    
  }
  
  }
  
  while(x2 == x){
    print(sprintf("%s, You win!", x2))
    break

  }
}
```
