---
title: 'Homework #1: SDGB-7844'
author: "Matthew Fein"
fordham_id_number: "A16797353"
fordham_email: "mfein@fordham.edu"
date: "10/19/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Each of the exercises below are from our textbook - _Introduction to Scientific Programming and Simulaiton Using R._ The exercises begin on p. 26.  Make sure to comment your code.  Points will be deducted for uncommented code!  

You should create new code blocks if necessary to answer the question.  You should also use plaintext to provide an explanation for your code if the question calls for one.

### Chapter 2 - Exercise 1a. 

```{r}z <- x^(a^b)          #x raiesed to the ath which has been raised to the bth


```

### Chapter 2 - Exercise 1b. 

```{r}z <- (x^a)^b            #x raised to the ath then raised to the bth


```

### Chapter 2 - Exercise 1c. 

```{r}z <- (3*x^3)+(2*x^2)+(6^x)+1
 

```

### Chapter 2 - Exercise 1d. 

```{r}
x <- 3.08                             #set number for division
z <- floor((x %% 1)*100)              # extract second digit after decimal
```

### Chapter 2 - Exercise 1e. 

```{r}
z <- z + 1              #increment z by 1

```


### Chapter 2 - Exercise 2a.

```{r}
x <- c(1:8,7:1)   #created vector ascending from 1-8 and descending from 7-1

```

### Chapter 2 - Exercise 2b.

```{r}
y <-rep(1:5, 1:5)     # create vector of 1-5 with 1-5 repeated 1-5 times each

```


### Chapter 2 - Exercise 2c. 

```{r}
A <- matrix(c(0,1,1,1,0,1,1,1,0), nrow = 3, ncol = 3)   #create matrix with                                                               #specific values

```

### Chapter 2 - Exercise 2d.

```{r}
B <- matrix(c(0,0,7,2,5,0,3,0,0), nrow = 3, ncol = 3)   #create matrix with                                                               #specific values

```

### Chapter 2 - Exercise 4

```{r}
x <- c(1:100)           #creates vector of alldigits from 1-100
y <- x[x%%2 != 0 & x%%3 !=0 & x%%7 != 0]    #creates vector of elements from x of 
                                            #which are not divisible by 2,3,7
```


### Chapter 2 - Exercise 5a. 

```{r}
queue <- c("Steve", "Russell", "Alison", "Liam") # create starting order
queue <- c(queue, "Barry")    # adds Barry to the last position of queue
```

### Chapter 2 - Exercise 5b. 

```{r}
queue <- queue[-1]    #removes the first person in line
```

### Chapter 2 - Exercise 5c.

```{r}
queue <- c("Pam",queue)   #adds Pam to the front of the line
```

### Chapter 2 - Exercise 5d. 

```{r}
queue <- queue[-5]        #removes Barry from last position
```

### Chapter 2 - Exercise 5e. 

```{r}
queue <- queue[which(queue != "Alison")] # return names which are not Alison

```

### Chapter 2 - Exercise 6

```{r}
rm(list=ls())                  #clear workspace
x <- 1                          # create X
print(x)                        # show X
x[3] <- 3                       #place 3 in the third position of x
print(x)                        # show X, NA in 2nd position
y <- c()                        #creates empty vector y
print(y)                        # shows y
y[2] <- 2                       # inserts 2 into the second position
print(y)                        # shows y
y[3] <- y[1]                    # inserts NA in the third position
print(y)                        # shows y
y[2] <- y[4]                    # inserts NA into the second position
print(y)                        # shows y
#z[1] <- 0                       # fails to run
```

### Chapter 2 - Exercise 7

```{r}
ID <- diag(10)                   # create identity matrix
ID<- ifelse(row(ID)==col(ID),ID,5)   #replace non diagonal elements with 5's

ID <- diag(10) 
ID[upper.tri(ID) | lower.tri(ID)] = 5 #alternate way of replacing 0s with 5s
```

